# manifoldflasso_jmlr
This repository contains code for generation of figures in the Manifold Coordinates with Physical Meaning’ (Koelle, Zhang, Meila, Chen).
The major steps are 1) configuring the virtual environment and 2) running the code.

1) Configuring the virtual environment
This part can be done prior to cloning the manifoldflasso_jmlr repository, and does not depend on being run in the same directory.

1.1) Create a virtual env with megaman dependencies

conda create -n manifold_env_resubmission python=3.5 -y
source activate manifold_env_resubmission
conda install --channel=conda-forge -y pip nose coverage cython numpy scipy scikit-learn pyflann pyamg h5py plotly

1.2) Install megaman

mkdir megaman_jmlr
cd megaman_jmlr
mkdir tmp
cd tmp/
git clone https://github.com/mmp2/megaman.git
cd megaman
python setup.py install
cd ../..

1.4) (Optional) configure jupiter notebook
conda install -c anaconda ipython
conda install -c anaconda ipykernel
ipython kernel install --user --name=manifold_env_resubmission --display-name=manifold_env_resubmission

2) Running the code

2.a) Toy Data)

The RigidEthanol and SwissRoll examples simulate data, and so can be run without adding other files.

2.b MDS Data)

The Toluene, Ethanol, and Malonaldehyde examples require paths to data to be specified.
 Data is available at xxx.

2.b.1 Agnostic dictionaries)

These will run much faster than full dictionaries, and are runnable on a 2015 macbook pro.

Results can be created e.g.

python codes/experiments/ethanol_1221_p12rep5n500_papernorm.py

or

sbatch codes/experiments/ethanol_1221_p12rep5n500_papernorm.sh

Figures can then be generated using aprioridictionaries_jmlrresubmission.ipynb

2.b.2 A priori dictionaries)

These are much slower.  1 replicate takes around 1 hour on a 2015 macbook pro. 

python codes/experiments/eth_1226_pallrep5n100.py

or

sbatch codes/experiments/eth_1226_pallrep5n100.sh



2.1) If running on a cluster
Use the .sh files e.g. 
sbatch swissroll_041420_nsel_50_nreps5.sh

2.2) If running locally
source activate manifold_env_resubmission





