\documentclass{article}
%\documentclass[]{article}
% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}

% to compile a preprint version, e.g., for submission to arXiv, add
% add the [preprint] option:
\usepackage[preprint]{jmlr2e}

\usepackage{natbib}
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2018}

%\usepackage[backslant]{aurical}       % Fontauri fonts 
\usepackage{bm}
\usepackage[utf8]{inputenc} % allow utf-8 input
%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{subfig}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{enumerate}
\usepackage{xcolor}
%\usepackage{graphicx}
\usepackage[font=small]{caption}
%\usepackage[labelformat = empty,position=top]{subcaption}
\usepackage[export]{adjustbox}
%\usepackage[]{algorithm2e}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{float}
\usepackage{placeins}
\usepackage{multirow}

\include{gradients-commands}
\newlength{\llw}
\setlength{\llw}{\linewidth}
\newlength{\picwi}

\begin{document}

%\title{A regression approach for explaining manifold embedding coordinates}
\title{Manifold Coordinates with Physical Meaning}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{\name Samson J. Koelle${}^1$ \email sjkoelle@uw.edu \\
       \name Hanyu Zhang${}^1$ \email hanyuz6@uw.edu \\
       \name Marina Meil\u{a}${}^{1,2}$ \email mmp@stat.washington.edu \\
       \name Yu-Chia Chen${}^2$ \email yuchaz@uw.edu \\
       \addr ${}^1$Department of Statistics\\
       University of Washington\\
       Seattle, WA 98195-4322, USA   \\
       ${}^2$Department of Electrical and Computer Engineering \\
       University of Washington \\
       Seattle, WA 98195, USA
       }
       
\editor{Francis Bach, David Blei, and Bernhard Sch{\"o}lkopf}
%\author{
%  Marina Meila\thanks{\texttt{www.stat.washington.edu/mmp}}
  %\And
 % Samson Koelle \\
  %Department of Statistics\\
 % University of Washington\\
 % Seattle, WA 98195-4322 \\
  %\texttt{\{mmp2,sjkoelle,hanyuz6\}@uw.edu} \\
  %\And
  %Hanyu Zhang
%  \And
%  Yu-Chia Chen\\
   %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
%}

\maketitle
\begin{abstract}
Manifold embedding algorithms map high-dimensional data down to
coordinates in a much lower-dimensional space. One of the aims of
dimension reduction is to find {\em intrinsic coordinates} that
describe the data manifold. The coordinates returned by the
embedding algorithm are abstract, and finding their physical or
domain-related meaning is not formalized and left to domain
experts. This paper studies the problem of recovering the
meaning of the new low-dimensional representation in a
 semi-automatic, principled fashion.  We propose a method to explain
embedding coordinates of a manifold as {\em non-linear} compositions
of functions from a user-defined dictionary. We show that this problem
can be set up as a sparse {\em linear Group Lasso} recovery problem,
find sufficient recovery conditions, and demonstrate its effectiveness
on data.

\end{abstract}

\begin{keywords}
Dimension reduction, manifold learning, functional regression, gradient, group lasso
\end{keywords}

\input{gradients-jmlr-intro}
\input{gradients-jmlr-problem}
\input{gradients-jmlr-background}
%\input{gradients-jmlr-flasso}
\input{gradients-jmlr-pullback}
\input{gradients-jmlr-related}
\input{gradients-jmlr-theory}
\input{gradients-jmlr-exp}
\input{gradients-jmlr-conclusion}

\medskip
\bibliographystyle{alpha}
\begin{small}
\bibliography{../mmp,../materials,../manifolds,../gradients}
\end{small}

\comment{original -proofs merged with unicity. we'll keep this file here because eventually some proofs will be move to the back. There is also a -supplement file that's not used...}
%\input{gradients-nips18-proofs}

\input{gradients-jmlr-appendix}

\end{document}
